\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1in]{geometry}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\newcommand{\qbin}[2]{\begin{bmatrix}#1\\#2\end{bmatrix}_q}
\title{Seed 8, Layer 2: The bracket tower $f_k^{(m)}$, exact $q$-monotonicity,\\ and the Divisibility Transfer Lemma}
\author{Seed 8 / Layer 2 / Round 2}
\date{2026-07-04}
\begin{document}
\maketitle

\section{Setup}

Fix $r=3$, level $d\not\equiv 0\pmod 3$, and a profile $c=(c_0,c_1,c_2)$ with
$c_0+c_1+c_2=d$. Let $g_m=g_m(q)$ denote the generating function for cylindric
partitions of profile $c$ with largest part exactly $m$ (so $g_0=1$), computed
by the profile-space transfer matrix. Set $h_m=(q;q)_m\,g_m$, and define the
tower of Round~1:
\[
D_0^m=h_m,\qquad D_k^m=D_{k-1}^m-q^k\,D_{k-1}^{m-1}\quad(1\le k\le m),\qquad
Q_n=D_n^n .
\]

\section{The bracket recursion}

Define $f_{-1}^{(m)}:=g_m$ and, for $k\ge 0$ and $m\ge k+1$,
\begin{equation}\label{eq:bracket}
f_k^{(m)} \;:=\; (1-q^{m-k})\,f_{k-1}^{(m)} \;-\; q^{k+1}\,f_{k-1}^{(m-1)} .
\end{equation}

\begin{lemma}[T1]\label{lem:T1}
For all $k\ge -1$ and $m\ge k+1$,
\[
D_{k+1}^{m} \;=\; (q;q)_{m-k-1}\; f_k^{(m)} .
\]
\end{lemma}

\begin{proof}
Induction on $k$. Base $k=-1$: $D_0^m=h_m=(q;q)_m\,g_m=(q;q)_m f_{-1}^{(m)}$
by definition. Inductive step: assume the claim at $k-1$ for all admissible
$m$. Then
\begin{align*}
D_{k+1}^m &= D_k^m - q^{k+1} D_k^{m-1}
   = (q;q)_{m-k}\,f_{k-1}^{(m)} - q^{k+1}(q;q)_{m-1-k}\,f_{k-1}^{(m-1)}\\
 &= (q;q)_{m-k-1}\Bigl[(1-q^{m-k})f_{k-1}^{(m)} - q^{k+1} f_{k-1}^{(m-1)}\Bigr]
   = (q;q)_{m-k-1}\, f_k^{(m)},
\end{align*}
using $(q;q)_{m-k}=(q;q)_{m-k-1}(1-q^{m-k})$ and recursion \eqref{eq:bracket}.
\end{proof}

\begin{corollary}[T2]
$Q_n = D_n^n = (q;q)_0\, f_{n-1}^{(n)} = f_{n-1}^{(n)}$.
\end{corollary}

Hence Warnaar's positivity conjecture is exactly the diagonal $m=k+1$ of the
bracket tower.

\begin{lemma}[Closed form]
For $k\ge -1$ and $m\ge k+1$,
\[
f_k^{(m)} \;=\; \sum_{i=0}^{k+1} (-1)^i\, q^{\binom{i+1}{2}}\,
   \qbin{k+1}{i}\, (q^{m-k};q)_{k+1-i}\; g_{m-i}.
\]
\end{lemma}

\begin{proof}
Combine Lemma~\ref{lem:T1} with the Round-1 closed form
$D_{k+1}^m=\sum_i(-1)^i q^{\binom{i+1}{2}}\qbin{k+1}{i}h_{m-i}$, divide by
$(q;q)_{m-k-1}$, and use $h_{m-i}=(q;q)_{m-i}g_{m-i}$ together with
$(q;q)_{m-i}/(q;q)_{m-k-1}=(q^{m-k};q)_{k+1-i}$ for $0\le i\le k+1$.
(Each quotient is a genuine polynomial, so the identity holds in
$\mathbb{Z}[[q]]$.)
\end{proof}

\section{The Master Conjecture (verified, not proved)}

\begin{conjecture}[Master]
For all $k\ge -1$, $m\ge k+1$, and $0\le j\le m-k-1$,
\[
(q;q)_j\, f_k^{(m)} \;\succeq\; 0
\]
(coefficientwise), and this is \emph{exact}: $(q;q)_{m-k}\,f_k^{(m)}$ has a
negative coefficient. In particular ($k=-1$): $g_m$ is exactly $m$-fold
$q$-monotone.
\end{conjecture}

The top case $j=m-k-1$ is $D_{k+1}^m\succeq 0$ (the Round-1 tower); the case
$(k,j)=(m-1,0)$ is $Q_m\succeq 0$. Verified computationally, with exactness at
every boundary, for: $d=2$ ($m\le 8$, $k\le 4$); $d=4$, all $5$ orbit
representatives ($m\le 7$, $k\le 4$); $d=5$, all $7$ representatives
($m\le 6$, $k\le 4$); $d=7$, all $6$ representatives ($m\le 5$, $k\le 3$).

Multiplying \eqref{eq:bracket} by $(q;q)_j$ shows the induction on $k$ needs,
at every level, the full ``monotonized domination''
$(q;q)_j(1-q^{m-k-1})f_k^{(m)}\succeq q^{k+2}(q;q)_j f_k^{(m-1)}$; positivity
alone does not chain. The one general reduction we can prove is:

\begin{lemma}[Divisibility Transfer]
$\dfrac{1-q^a}{1-q^b}$ has nonnegative coefficients iff $b\mid a$.
Consequently, if $F\in\mathbb{Z}[[q]]$ satisfies $(q;q)_J F\succeq 0$ and
$S\subset\mathbb{Z}_{>0}$ is a finite multiset admitting an injection
$\sigma\colon S\to\{1,\dots,J\}$ with $\sigma(a)\mid a$ for all $a\in S$, then
$\prod_{a\in S}(1-q^a)\,F\succeq 0$.
\end{lemma}

\begin{proof}
If $b\mid a$, $\frac{1-q^a}{1-q^b}=1+q^b+\dots+q^{a-b}$. If $b\nmid a$, the
series $\frac{1-q^a}{1-q^b}=(1-q^a)\sum_{t\ge0}q^{bt}$ has coefficient $-1$ at
the least exponent of the form $a+bt$ not divisible by $b$ (namely $a$ itself
if $b\nmid a$; indeed the coefficient of $q^n$ is $\#\{t:bt=n\}-\#\{t:a+bt=n\}$,
which is $-1$ at $n=a$ when $b\nmid a$). For the consequence, write
\[
\prod_{a\in S}(1-q^a)\,F
=\Bigl[\prod_{a\in S}\frac{1-q^a}{1-q^{\sigma(a)}}\Bigr]
 \Bigl[\prod_{b\in\{1..J\}\setminus\sigma(S)}\frac{1}{1-q^{b}}\Bigr]
 \,(q;q)_J F,
\]
a product of three coefficientwise-nonnegative series.
\end{proof}

Empirically (29{,}482 subsets tested over four $(d,c,m,k)$ cases, zero
violations) the divisibility-matching condition also appears close to
necessary. However, for the induction step one needs
$\{1,\dots,j\}\cup\{m-k-1\}$ to match into $\{1,\dots,j+1\}$, which forces
$(j+1)\mid(m-k-1)$: the Transfer Lemma alone proves only these arithmetic
slices, not the full tower.

\section{The chain model and two structural lemmas}

Cylindric partitions of profile $c$ with largest part $\le m$ biject
(conjugate model, verified against the transfer matrix) with chains
$a^{(1)}\ge\dots\ge a^{(m)}$ in
$S=\{a\in\mathbb{Z}_{\ge0}^3 : a_i\le a_{i-1}+c_i \text{ cyclically}\}$,
weighted by $\sum_s |a^{(s)}|$; largest part exactly $m$ iff all levels are
nonzero.

\begin{lemma}[Slack]
Every $a\in S$ has a \emph{slack} coordinate $i$ (i.e.\ $a_i<a_{i-1}+c_i$),
since $d=\sum c_i\ge 1$ makes all three constraints tight impossible.
\end{lemma}

\begin{lemma}[Switch]
If $i$ is tight at $a$ and slack at $b$, with $a\ge b$ componentwise
($a,b\in S$), then $a_i=a_{i-1}+c_i\ge b_{i-1}+c_i>b_i$, so $a_i>b_i$ strictly.
\end{lemma}

These make box-adding ribbon maps well defined, but three distinct designs of
a weight-$(+m)$ injection $\psi\colon C_m\to C_m$ (needed for
$f_0^{(m)}\succeq 0$) failed: a greedy least-slack ribbon is total but not
injective; a claim-based inverse-by-construction map is injective but not
total (orphan states exist for every profile); a top-level add reduces
circularly to $(1-q)g_m\succeq 0$. The pattern of failure suggests a global
canonical structure --- plausibly a crystal operator on chains (cf.\ Tingley's
question on lifting $U_q(\widehat{\mathfrak{sl}}_n)$ crystal structures to
cylindric plane partitions) --- is required.

\subsection*{Key example}
$d=4$, $c=(2,1,1)$, $m=2$: $f_0^{(2)}=(1-q^2)g_2-q\,g_1$ has expansion
$3q^3+7q^4+10q^5+13q^6+\cdots\succeq 0$, and
$(1-q)f_0^{(2)}=D_1^2=3q^3+4q^4+3q^5+3q^6+2q^7+2q^8+q^9+q^{10}+q^{12}
\succeq 0$, while $(1-q)(1-q^2)f_0^{(2)}$ has coefficient $-1$ at $q^6$:
exact $1$-fold monotonicity, matching Master at $(k,m)=(0,2)$.

\end{document}
